---
title: "Assignment9_MA"
author: "Akshita"
date: "2024-04-18"
output: html_document
---
1. Model Development
```{r}
library(ggplot2)
library(cowplot)
library(pROC)

fish_data <- read.csv("C:/Users/HP/Desktop/USA/Spring'24/Multivariate/Datasets/Fish_dataset.csv")
str(fish_data)

# Subset the fish data to create a binary classification task
binary_fish_data <- subset(fish_data, Species == "Bream" | Species != "Bream")

# Convert the target variable to a binary outcome (1 for Bream, 0 for non-Bream)
binary_fish_data$IsBream <- ifelse(binary_fish_data$Species == "Bream", 1, 0)

# Fit logistic regression model
binary_fish_model <- glm(IsBream ~ ., data = binary_fish_data, family = "binomial")

# Summary of the model
summary(binary_fish_model)

```
Overall, the output suggests that none of the predictor variables are statistically significant in predicting whether a fish is a Bream or not, as indicated by the high p-values. Additionally, the extremely small residual deviance indicates an excellent fit of the model to the data. However, the lack of significant predictors may raise concerns about the model's usefulness in accurately predicting the outcome. 


2. Model Acceptance & 3. Residual Analysis
```{r}
# Goodness of Fit
# Compare null deviance and residual deviance
null_deviance <- summary(binary_fish_model)$null.deviance
residual_deviance <- summary(binary_fish_model)$deviance
deviance_reduction <- null_deviance - residual_deviance
print(paste("Null Deviance:", null_deviance))
print(paste("Residual Deviance:", residual_deviance))
print(paste("Deviance Reduction:", deviance_reduction))

# AIC
AIC_value <- AIC(binary_fish_model)
print(AIC_value)

# Assumptions Checking (Diagnostic Plots)
par(mfrow = c(2, 2))  # Set up a 2x2 plot layout
plot(binary_fish_model)
```
Deviance Reduction:
The deviance reduction from the null model to the residual model is substantial, indicating that the model explains a significant portion of the variability in the data.

AIC:
AIC value of 26 suggests a good balance between model fit and complexity. Lower AIC values are preferred, indicating better model performance.

Deviance:
Residual deviance close to zero indicates that the model fits the data well, with little unexplained variability.

4. Prediction 
```{r}
# Predict probabilities for each observation
predicted_probabilities <- predict(binary_fish_model, type = "response", newdata = binary_fish_data)

# View the predicted probabilities
head(predicted_probabilities)

# Set a threshold probability (e.g., 0.5)
threshold <- 0.5

# Convert predicted probabilities to binary predictions based on the threshold
predicted_classes <- ifelse(predicted_probabilities >= threshold, 1, 0)

# Calculate accuracy
accuracy <- mean(predicted_classes == binary_fish_data$IsBream)
print(paste("Accuracy:", accuracy))

# Create a data frame with actual and predicted values
prediction_df <- data.frame(
  Actual = binary_fish_data$IsBream,
  Predicted_Probability = predicted_probabilities)

# Sort the data frame by predicted probabilities for better visualization
prediction_df <- prediction_df[order(prediction_df$Predicted_Probability), ]

# Create a sequence of observation numbers
observation_numbers <- seq_along(predicted_probabilities)

# Plot predicted probabilities
plot(observation_numbers, predicted_probabilities, type = "l", col = "blue",
     xlab = "Observation", ylab = "Probability", 
     main = "Predicted Probabilities vs. Actual Outcomes")

# Add actual outcomes as points (red for Bream, green for Non-Bream)
points(observation_numbers[binary_fish_data$IsBream == 1], 
       rep(1, sum(binary_fish_data$IsBream == 1)), col = "red", pch = 19)
points(observation_numbers[binary_fish_data$IsBream == 0], 
       rep(0, sum(binary_fish_data$IsBream == 0)), col = "green", pch = 19)

# Add legend
legend("topright", legend = c("Predicted", "Actual (Bream)", "Actual (Non-Bream)"),
       col = c("blue", "red", "green"), lty = 1, pch = c(NA, 19, 19), cex = 0.8)

```


5. Model Accuracy
```{r}
# Predicted probabilities from the logistic regression model
predicted_probabilities <- predict(binary_fish_model, type = "response")

# Convert predicted probabilities to binary predictions (1 for Bream, 0 for Non-Bream)
predicted_classes <- ifelse(predicted_probabilities >= 0.5, 1, 0)

# Calculate accuracy
accuracy <- mean(predicted_classes == binary_fish_data$IsBream)
cat("Model Accuracy:", accuracy, "\n")

# Predict classes using the logistic regression model
predicted_classes <- ifelse(predict(binary_fish_model, newdata = binary_fish_data, type = "response") > 0.5, "Bream", "Not Bream")

# Create a confusion matrix
confusion_matrix <- table(predicted_classes, as.factor(binary_fish_data$IsBream))

# Print confusion matrix
print(confusion_matrix)

# Calculate ROC curve
roc_curve <- roc(binary_fish_data$IsBream, predicted_probabilities)

# Plot ROC curve
plot(roc_curve, main = "ROC Curve for Logistic Regression Model",
     col = "blue", lwd = 2)

# Add diagonal line for reference
abline(a = 0, b = 1, col = "red", lty = 2)

# Add legend
legend("bottomright", legend = c("ROC Curve", "Random Guessing"),
       col = c("blue", "red"), lty = 1:2, cex = 0.8)

```