---
title: "Assignment6_MA"
author: "Akshita"
date: "2024-03-24"
output: html_document
---
1. Decide how many Factors are ideal for your dataset.
```{r}
library(psych)

fish_data <- read.csv("C:/Users/HP/Desktop/USA/Spring'24/Multivariate/Datasets/Fish_dataset.csv")

attach(fish_data)
fish_data[1]
fit.pc_fish <- principal(fish_data[-1], nfactors=4, rotate="varimax")
fit.pc_fish

# Extract eigenvalues
eigenvalues <- fit.pc_fish$values
# Create scree plot
plot(1:length(eigenvalues), eigenvalues, type="b", 
      xlab="Component Number", ylab="Eigenvalues", main="Scree Plot")
# Add a horizontal line at y = 1 for reference (Kaiser's criterion)
abline(h=1, col="red", lty=2)
# Add text label for Kaiser's criterion
text(length(eigenvalues), 1, "Kaiser's Criterion (Eigenvalue = 1)", pos=4, col="red")

round(fit.pc_fish$values, 3)
fit.pc_fish$loadings

# Loadings with more digits
for (i in c(1,2,3,4)) { print(fit.pc_fish$loadings[[1,i]])}
# Communalities
fit.pc_fish$communality
# Rotated factor scores, Notice the columns ordering: RC1, RC3, RC2 and RC4
fit.pc_fish$scores
```

According to the scree plot, the first factor (RC1) explains a significant portion of the variance (3.35), followed by a substantial drop for the second factor (1.60). This suggests that one factor might be sufficient.Taking RC2 also in consideration, the model might be a little complex but can give improve variablity for the dataset, hence we can consider 2 factors for this dataset.

2. Explain the output for your factor model
```{r}
#FA model
fa.plot(fit.pc_fish)
fa.diagram(fit.pc_fish)
vss(fish_data[-1])

# Computing Correlation Matrix
corrm.fish <- cor(fish_data[-1])
corrm.fish
plot(corrm.fish)

fish_pca <- prcomp(fish_data[-1], scale=TRUE)
summary(fish_pca)
plot(fish_pca)

# A table containing eigenvalues and %'s accounted, follows. Eigenvalues are the sdev^2
(eigen_fish <- round(fish_pca$sdev^2,3))
round(fit.pc_fish$values, 3)
names(eigen_fish) <- paste("PC",1:6,sep="")
eigen_fish

sumlambdas <- sum(eigen_fish)
sumlambdas

propvar <- round(eigen_fish/sumlambdas,2)
propvar

cumvar_fish <- cumsum(propvar)
cumvar_fish

matlambdas <- rbind(eigen_fish,propvar,cumvar_fish)
matlambdas

rownames(matlambdas) <- c("Eigenvalues","Prop. variance","Cum. prop. variance")
rownames(matlambdas)

eigvec.fish <- fish_pca$rotation
print(fish_pca)

# Taking the first two PCs to generate linear combinations for all the variables with two factors
pcafactors.fish <- eigvec.fish[,1:2]
pcafactors.fish

# Multiplying each column of the eigenvector’s matrix by the square-root of the corresponding eigenvalue in order to get the factor loadings
unrot.fact.fish <- sweep(pcafactors.fish,MARGIN=2,fish_pca$sdev[1:2],`*`)
unrot.fact.fish

# Computing communalities
communalities.fish <- rowSums(unrot.fact.fish^2)
communalities.fish

# Performing the varimax rotation. The default in the varimax function is norm=TRUE thus, Kaiser normalization is carried out
rot.fact.fish <- varimax(unrot.fact.fish)
#View(unrot.fact.fish)
rot.fact.fish
```

Loadings: For PC1, Weight, Length1, Length2 and Length3 have high loadings which indicate that they contribute significantly to PC1.

SS Loadings: It indicates the amount of variance explained by each PC. For instance, PC1 explains 3.867 units of variance, while PC2 explains 1.913 units.

Proportion Variance: It represents the percentage of total variance explained by each PC. Here, PC1 explains 64.5% of the total variance, while PC2 explains 31.9%.

Cummulative Varicance: It indicates how much of the total variance is explained cumulatively by each PC. Here, after PC1 and PC2, 96.3% of the total variance is explained.

3. Show the columns that go into each factor.
```{r}
# The print method of varimax omits loadings less than abs(0.1). In order to display all the loadings, it is necessary to ask explicitly the contents of the object $loadings
fact.load.fish <- rot.fact.fish$loadings[1:6,1:2]
fact.load.fish

scale.fish <- scale(fish_data[-1])
scale.fish
as.matrix(scale.fish)%*%fact.load.fish%*%solve(t(fact.load.fish)%*%fact.load.fish)

# Show the columns (variables) that go into each factor
for (i in 1:ncol(fact.load.fish)) {
  cat("Variables for Factor PC", i, ":\n")
  print(rownames(fact.load.fish)[order(abs(fact.load.fish[, i]), decreasing = TRUE)])
  cat("\n")
}
```

For PC1, "Length1" "Length2" "Length3" "Weight" are the primary influence.

For PC2, "Height"  "Width"   "Weight"  "Length3" are the primary influence.

4. Perform some visualizations using the factors.
```{r}

# Convert factor loadings matrix to a data frame
fact_loadings_df <- as.data.frame(fact.load.fish)
# Add a column for variable names
fact_loadings_df$Variable <- rownames(fact_loadings_df)
# Reshape data for plotting
fact_loadings_df_long <- reshape2::melt(fact_loadings_df, id.vars = "Variable", variable.name = "Factor")

library(ggplot2)
#HeatMap
ggplot(fact_loadings_df_long, aes(x = Factor, y = Variable, fill = value)) +
     geom_tile() +
     scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
     labs(x = "Factor", y = "Variable", title = "Factor Loadings Heatmap") +
     theme_minimal() +
     theme(axis.text.x = element_text(angle = 45, hjust = 1))

#BarPlot
ggplot(fact_loadings_df_long, aes(x = Factor, y = value, fill = Variable)) +
     geom_bar(stat = "identity", position = "dodge") +
     labs(x = "Factor", y = "Factor Loading", fill = "Variable", title = "Factor Loadings Bar Plot") + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1),
           legend.position = "right")

